Overlapping Batches for the Assessment of Solution Quality in Stochastic Programs

### 2014-07-05 Referee1_Response.tex

\begin{itemize}
\item[12.] \textit{ p14, l23: Please add $j = 1$ to the sum.}
\end{itemize}

\noindent 
Added. Thanks for catching this.
\medskip

\begin{itemize}
\item[13.] \textit{p14, l35: The authors have gone to some length to distinguish different types of $A_j$ terms.
However, this equation has an $\mathbb{E}A_j^2$ term that has been collapsed from a sum, but we don't know ``which $j$''. 
Should there be a $\max_j \mathbb{E}A_j^2$ or some other treatment here?}
\end{itemize}

\noindent 
This is because the center batches are identically distributed. 
Therefore, $\mathbb{E}A_j^2$ for all the center batches $j=\lceil \bar{\gamma}^{-1} \rceil, \ldots,n_b - \lceil \bar{\gamma}^{-1} \rceil$ are the same. 
To make this clear, we introduced the notation $A_j^c$ for center batches and $A_j^f$ and $A_j^e$ for the fringe batches that appear at the front and end of each set of nonoverlapping batches, respectively. 
We also emphasized that each front fringe batch has an identically distributed pair at the end fringes, and that all the center batches are identically distributed. 
%Thanks for catching this. 
(Please note that we added brackets to expectations to improve readability as suggested by the other referee.)
\medskip

### 2014-07-05 Referee2_Response.tex

\noindent 
xxx  We agree. 
We changed the introduction to clearly state the contributions of the paper and \ldots  and the presentation of the paper significantly improved.  Thank you.
\medskip

\begin{itemize}
\item[2.] \textit{It is not necessary to go into such detail in the OBM review.}
\end{itemize}

\smallskip

\noindent 
We shortened and streamlined the OBM review, focusing on the main results used in the paper, by combining the previous two sections---\S 2.1 (Overlapping Batch Means) and \S 3.1 (Variably Overlapping Batches)---along with other explanations on OBM that appeared later in the paper. 
The combined review is in the background section, \S 2.1 (Batch Means). 
To improve readability, we created two subsections: \S 2.1.1 (Nonoverlapping Batch Means) and \S 2.1.2 (Overlapping Batch Means). 
While we removed a number of equations and shortened the text, we kept some explanations that are necessary for a reader who is not familiar with OBM (e.g., readers with an optimization focus) to follow the main ideas. 
These basic explanations appear, for instance, in the first paragraph of \S 2.1.2 (Overlapping Batch Means).  
We hope to be able to reach a broader audience this way. 
As a result of these changes, 
\begin{itemize}
	\item[(i)]  instead of former equations (2) and (5), both on overlapping batch means variance estimator, we have only one equation (2);

### 2014-07-05 omrp_lb_rev.tex
Referee 1, Minor Comment 13

The amount of asymptotic reduction in $\var{\vartg{\ybb}}$ for a given $0\leq \gamma< m$ relative its nonoverlapping counterpart 
%$\var{\vtm{\ybb}}$ 
depends on the asymptotic ratio of $\gamma/m$, which we denote by $\gammab$. 
The maximally-overlapping case ($\gamma = 1$) corresponds to $\gammab=0$, and the nonoverlapping case corresponds to $\gammab=1$. 
\cite{damerdji1994strong,damerdji1995mean} studied batching structures that guarantee almost sure and mean-square consistency of $\vartg{\ybb}$ in the nonoverlapping and maximally-overlapping cases. 
Under appropriate conditions, these results, along with those of \cite{Meketon1984} and \cite{Welch1987}, establish that 
\begin{equation}
\label{eq:varvarred}
\frac{n}{m}\var{n \vartg{\ybb}} \rightarrow \beta_{\gammab} \sigma^4
\end{equation}
as the batch size $m$ and the number of batches $n/m$ tend to infinity. 
In (\ref{eq:varvarred}), $\beta_1 = 2$ for nonoverlapping batches, $\beta_0 = \frac{4}{3}$ for maximally overlapping batches, and more generally, $\beta_{\gammab}$ is twice the value of 
\begin{equation} \label{eq:var_reduct_formula}
	\gammab \left( 1 + 2 \sum_{j=1}^{\cogb-1} (1-j\gammab)^2 \right),
\end{equation}
where $\lceil \cdot \rceil$ denotes the smallest integer that is greater than or equal to its argument.
If $\ogb$ is an integer, (\ref{eq:var_reduct_formula}) simplifies to $\frac{2+\gammab^2}{3}$.
This means that the maximally-overlapping variance estimator has a variance that is two-thirds ($66.67\%$) of the nonoverlapping estimator.
As the amount of overlap decreases, the variance increases. 
When only half of the observations overlap ($\gammab = 1/2$ or $\gamma=m/2$), for instance, the variance is 75\% of the nonoverlapping counterpart.

@@

	The fringe batches come in identically distributed pairs.  
        The first and last batch are identically distributed, as are the second and second-to-last, etc.  
        We know this because we could reverse the order of the i.i.d.\ sample and perform the batching again to get the exact same batch structure.  
        Rewriting to take this into account, letting $A_j = (\gb_j - \gbb)^2 - (\db_j - \dbb)^2$, and continuing from (\ref{eq:vvar_conv_pf_1}),
	\begin{align*}
		& \left( \frac{n}{dt} \right)^2 n_b \e{ \sum_{j=1}^{n_b} \left[ (\gb_j - \gbb)^2 - (\db_j - \dbb)^2 \right]^2 } \\
		 =  & \left( \frac{n}{dt} \right)^2 n_b \e{\sum_{j=1}^{\left\lceil \gammab^{-1} \right\rceil} \left( A_j \right)^2 \
		 + \sum_{j=\left\lceil \gammab^{-1} \right\rceil+1}^{n_b-\left\lceil \gammab^{-1} \right\rceil} \left( A_j \right)^2 \
		 + \sum_{j=n_b-\left\lceil \gammab^{-1} \right\rceil+1}^{n_b} \left( A_j \right)^2 
		} \\
		 = & \left( \frac{n}{dt} \right)^2 n_b \sum_{j=\left\lceil \gammab^{-1} \right\rceil+1}^{n_b-\left\lceil \gammab^{-1} \right\rceil} \e{ \left( A_j \right)^2 } + \left( \frac{n}{dt} \right)^2 (n_b) 2\sum_{j=1}^{\left\lceil \gammab^{-1} \right\rceil} \e{ \left( A_j \right)^2 } \\
		 = & \left( \frac{n}{dt} \right)^2 n_b \left(n_b - 2\left\lceil \gammab^{-1} \right\rceil \right) \e{ \left( A_j \right)^2 }
		+ \left( \frac{n}{dt} \right)^2 (n_b) 2\sum_{j=1}^{\left\lceil \gammab^{-1} \right\rceil} \e{ \left( A_j \right)^2 }.
	\end{align*}	
	We have $\left\lceil \gammab^{-1} \right\rceil + 1$ expectation terms, each of which has a coefficient that is less than $\left(\frac{n}{dt}\right)^2n_b^2 < (2m)^2$.  
        So each expectation can be bounded by
	\begin{align}
		\left( 2m \right)^2 \e{ \left( A_j \right)^2 } & = \left( 2m \right)^2 \e{ \left((\gb_j - \gbb) - (\db_j - \dbb)\right)^2 \left((\gb_j - \gbb) + (\db_j - \dbb)\right)^2 } \notag \\
		& \leq \left( 2m \right)^2 \sqrt{\e{\left((\gb_j - \gbb) - (\db_j - \dbb)\right)^4}} \sqrt{\e{\left((\gb_j - \gbb) + (\db_j - \dbb)\right)^4 }} \label{eq:var_reduct_cauchy2} \\
		& \leq \left( 2m \right)^2 \sqrt{\e{\left((\gb_j - \gbb) - (\db_j - \dbb)\right)^4}} \sqrt{8 \left(\e{(\gb_j - \gbb)^4} + \e{(\db_j - \dbb)^4} \right)}, \label{eq:var_reduct_holder2}
	\end{align}
	where equation (\ref{eq:var_reduct_cauchy2}) follows from the Cauchy-Schwartz inequality, and equation (\ref{eq:var_reduct_holder2}) follows from H\"{o}lder's inequality.  
        Since we can bound both expectation terms in (\ref{eq:var_reduct_holder2}) by A2, we can demonstrate $\sqrt{\frac{n}{m}}(nVG - nVD) \xrightarrow{L^2} 0$ by focusing on only one of the terms.  
        Focusing on the first expectation term in (\ref{eq:var_reduct_holder2}) and returning the $\tfrac{n}{m}$, by H\"{o}lder's inequality,
	\begin{align*}
		\frac{n}{m} 4m^2 \e{\left((\gb_j - \gbb) - (\db_j - \dbb)\right)^4}	& \leq 32 \frac{n}{m} m^2\left( \e{(\db_j - \gb_j)^4} + \e{(\dbb - \gbb)^4} \right).
	\end{align*}	
	By Lemma \ref{lem:gb_gbb_l4}, both expectation terms above converge to zero with exponential rates.  
        With a batching structure $m_l = \lfloor n_l^r \rfloor$ for some $0<r<1$, we see that $\frac{n_l}{m_l} \e{(n_l VG_l - n_l VD_l)^2} \rightarrow 0$.
\end{proof}

Theorem \ref{thm:varvar_conv} gives conditions under which the same asymptotic lower variances are achieved for the OMRP optimality gap variance estimator relative to its nonoverlapping counterpart MRP. 
A hypothesis of Theorem \ref{thm:varvar_conv} is a batching structure that ensures convergence of $\frac{n_l}{m_l}\var{n_l VD_l}$ to the appropriate $\beta_{\gammab} \sigma^4_{\xh}$, ensuring the mean-square consistency of $n_l VD_l$. 
\citet{damerdji1995mean} shows that mean-square convergence is guaranteed by $\tfrac{1}{2} + \tfrac{1}{4+\epsilon} < r < 1$ for both nonoverlapping and maximally overlapping batches under the batching structure $m_l = \lfloor n_l^r \rfloor$ and $4+\epsilon$ finite moments assumption A2$^\prime$.  
With A2, the range of $r$ that ensures mean-square consistency can be found by setting $\epsilon=0$.

%We will empirically examine the variance reduction for small sample sizes in \S \ref{sec:comp}.

>>>>

The amount of asymptotic reduction in $\var{\vartg{\ybb}}$ for a given $0\leq \gamma< m$ relative its nonoverlapping counterpart 
%$\var{\vtm{\ybb}}$ 
depends on the asymptotic ratio of $\gamma/m$, which we denote by $\gammab$. 
The maximally-overlapping case ($\gamma = 1$) corresponds to $\gammab=0$, and the nonoverlapping case corresponds to $\gammab=1$. 
\cite{damerdji1994strong,damerdji1995mean} studied batching structures that guarantee almost sure and mean-square consistency of $\vartg{\ybb}$ in the nonoverlapping and maximally-overlapping cases. 
Under appropriate conditions, these results, along with those of \cite{Meketon1984} and \cite{Welch1987}, establish that 
\begin{equation}
\label{eq:varvarred}
\frac{n}{m}\var{n \vartg{\ybb}} \rightarrow \beta_{\gammab} \sigma^4
\end{equation}
as the batch size $m$ and then the number of batches $n/m$ tend to infinity. 
In (\ref{eq:varvarred}), $\beta_1 = 2$ for nonoverlapping batches, $\beta_0 = \frac{4}{3}$ for maximally overlapping batches, and more generally, $\beta_{\gammab}$ is twice the value of 
\begin{equation} \label{eq:var_reduct_formula}
	\gammab \left( 1 + 2 \sum_{j=1}^{\cogb-1} (1-j\gammab)^2 \right),
\end{equation}
where $\lceil \cdot \rceil$ denotes the smallest integer that is greater than or equal to its argument.
If $\ogb$ is an integer, (\ref{eq:var_reduct_formula}) simplifies to $\frac{2+\gammab^2}{3}$.
This means that the maximally-overlapping variance estimator has a variance that is two-thirds ($66.67\%$) of the nonoverlapping estimator.
As the amount of overlap decreases, the variance increases. 
When only half of the observations overlap ($\gammab = 1/2$ or $\gamma=m/2$), for instance, the variance is 75\% of the nonoverlapping counterpart.

@@

	The fringe batches come in identically distributed pairs.  
        The first and last batch are identically distributed, as are the second and second-to-last, etc.  
        We know this because we could reverse the order of the i.i.d.\ sample and perform the batching again to get the exact same batch structure.  
		  The center batches are identically distributed.  
        Let $A_j = (\gb_j - \gbb)^2 - (\db_j - \dbb)^2$.
		  Furthermore, let $A_j^f$, $j=1,\ldots,\left\lceil \gammab^{-1} \right\rceil$ denote the fringe batches at the front of each set of nonoverlapping batches, $A_j^c$, $j=\left\lceil \gammab^{-1} \right\rceil+1, \ldots n_b - \left\lceil \gammab^{-1} \right\rceil$ denote the center batches, and let $A_j^e$, $j=n_b - \left\lceil \gammab^{-1} \right\rceil+1,\ldots,n_b$ denote the fringe batches at the end of each set of nonoverlapping batches. 
		  Continuing from (\ref{eq:vvar_conv_pf_1}),
	\begin{align*}
		& \left( \frac{n}{dt} \right)^2 n_b \e{ \sum_{j=1}^{n_b} \left[ (\gb_j - \gbb)^2 - (\db_j - \dbb)^2 \right]^2 } \\
		 =  & \left( \frac{n}{dt} \right)^2 n_b \e{\sum_{j=1}^{\left\lceil \gammab^{-1} \right\rceil} \left( A_j^f \right)^2 \
		 + \sum_{j=\left\lceil \gammab^{-1} \right\rceil+1}^{n_b-\left\lceil \gammab^{-1} \right\rceil} \left( A_j^c \right)^2 \
		 + \sum_{j=n_b-\left\lceil \gammab^{-1} \right\rceil+1}^{n_b} \left( A_j^e \right)^2 
		} \\
		 = & \left( \frac{n}{dt} \right)^2 n_b \sum_{j=\left\lceil \gammab^{-1} \right\rceil+1}^{n_b-\left\lceil \gammab^{-1} \right\rceil} \e{ \left( A_j^c \right)^2 } + \left( \frac{n}{dt} \right)^2 (n_b) 2\sum_{j=1}^{\left\lceil \gammab^{-1} \right\rceil} \e{ \left( A_j^f \right)^2 } \\
		 = & \left( \frac{n}{dt} \right)^2 n_b \left(n_b - 2\left\lceil \gammab^{-1} \right\rceil \right) \e{ \left( A_j^c \right)^2 }
		+ \left( \frac{n}{dt} \right)^2 (n_b) 2\sum_{j=1}^{\left\lceil \gammab^{-1} \right\rceil} \e{ \left( A_j^f \right)^2 }.
	\end{align*}	
	These equalities follow because each front fringe batch $\e{(A_j^f)^2}$ has an identical pair in the end fringe batch $\e{(A_j^e)^2}$ and all center batches have identical values of $\e{(A_j^c)^2}$.  
	As a result, we have $\left\lceil \gammab^{-1} \right\rceil + 1$ expectation terms, each of which has a coefficient that is less than $\left(\frac{n}{dt}\right)^2n_b^2 < (2m)^2$.  
        Removing the notation for fringe or center batch for simplicity, each expectation can be bounded by
	\begin{align}
		\left( 2m \right)^2 \e{ \left( A_j \right)^2 } & = \left( 2m \right)^2 \e{ \left((\gb_j - \gbb) - (\db_j - \dbb)\right)^2 \left((\gb_j - \gbb) + (\db_j - \dbb)\right)^2 } \notag \\
		& \leq \left( 2m \right)^2 \sqrt{\e{\left((\gb_j - \gbb) - (\db_j - \dbb)\right)^4}} \sqrt{\e{\left((\gb_j - \gbb) + (\db_j - \dbb)\right)^4 }} \label{eq:var_reduct_cauchy2} \\
		& \leq \left( 2m \right)^2 \sqrt{\e{\left((\gb_j - \gbb) - (\db_j - \dbb)\right)^4}} \sqrt{8 \left(\e{(\gb_j - \gbb)^4} + \e{(\db_j - \dbb)^4} \right)}, \label{eq:var_reduct_holder2}
	\end{align}
	where equation (\ref{eq:var_reduct_cauchy2}) follows from the Cauchy-Schwartz inequality, and equation (\ref{eq:var_reduct_holder2}) follows from H\"{o}lder's inequality.  
        Since we can bound both expectation terms in (\ref{eq:var_reduct_holder2}) by A2, we can demonstrate $\sqrt{\frac{n}{m}}(nVG - nVD) \xrightarrow{L^2} 0$ by focusing on only one of the terms.  
        Focusing on the first expectation term in (\ref{eq:var_reduct_holder2}) and returning the $\tfrac{n}{m}$, by H\"{o}lder's inequality,
	\begin{align*}
		\left(\frac{n}{m} 4m^2\right)^2 \e{\left((\gb_j - \gbb) - (\db_j - \dbb)\right)^4}	& \leq 8 \left( \frac{n}{m} 4m^2\right)^2 \left( \e{(\db_j - \gb_j)^4} + \e{(\dbb - \gbb)^4} \right).
	\end{align*}	
	By Lemma \ref{lem:gb_gbb_l4}, both expectation terms above converge to zero with exponential rates.  
        With a batching structure $m_l = \lfloor n_l^r \rfloor$ for some $0<r<1$, we see that $\frac{n_l}{m_l} \e{(n_l VG_l - n_l VD_l)^2} \rightarrow 0$.
\end{proof}

Theorem \ref{thm:varvar_conv} gives conditions under which the same asymptotically lower variances are achieved for the OMRP optimality gap variance estimator relative to its nonoverlapping counterpart MRP. 
We will empirically examine the variance of OMRP estimator for small sample sizes in \S \ref{sec:comp}. 
A hypothesis of Theorem \ref{thm:varvar_conv} is a batching structure that ensures convergence of $\frac{n_l}{m_l}\var{n_l VD_l}$ to the appropriate $\beta_{\gammab} \sigma^4_{\xh}$, ensuring the mean-square consistency of $n_l VD_l$. 
\citet{damerdji1995mean} shows that mean-square convergence is guaranteed by $\tfrac{1}{2} + \tfrac{1}{4+\epsilon} < r < 1$ for both nonoverlapping and maximally overlapping batches under the batching structure $m_l = \lfloor n_l^r \rfloor$ and $4+\epsilon$ finite moments assumption A2$^\prime$.  
With A2, the range of $r$ that ensures mean-square consistency can be found by setting $\epsilon=0$.
